---
title: "A Simple Blockchain"
author: "Stephanie Jordan"
date: "`r Sys.Date()`"
#output: rmarkdown::pdf_document
output: rmarkdown::html_document

---

```{r setup, include = FALSE}
library(knitr)
library(digest)
knitr::opts_chunk$set(collapse = TRUE)
```

## Description
In this project, we will create a simple "mini" blockchain in R (without mining). 
This project is based on examples with R:

+ [Understanding Blockchains by Coding One in R](https://www.datacamp.com/community/tutorials/blockchain-r)
+ [Building a blockchain in R](http://users.dimi.uniud.it/~massimo.franceschet/HEX0x6C/blockchain/blockchainR.html )

### Data

In our example, the data for the blockchain must be in form of a *list*.
E.g. 

```{r Data, echo=TRUE}
NewData      <- list("Jordan",matrix(1:4,ncol=2)
                     ,hist(rnorm(234),plot=FALSE),200)
```

In this example, the data "NewData" is very heterogeneous - the list of length `r length(NewData) ` consists of different classes. 

### Structure of a block

A block is a **list**, which consits of 6 elements (in our simple case):

+ **index** index of the block
+ **timestamp** timestamp of the block
+ **data** data to be stored in the block
+ **previous_hash** previous hash (whereas for the genesis Block it is set to 0)
+ **proof proof** (provided by Proof of Work function)
+ **new_hash** hash of current block

### Hash function

Function that creates the "hashed block". For our hashes, we use the Sha-256 algorithm as implemented in the package **digest**

``` {r, echo= TRUE}
  hash_block <- function(block){
  # block: block to be hashed

    block$new_hash <- digest(c(block$index,
                             block$timestamp,
                             block$data,
                             block$previous_hash), "sha256")
  return(block)
}
```
### Proof of work functions

A new block needs to be validated and approved. We provide two different proof of work (PoW) functions:

#### Function **"proof_of_work"**

Increment the proof number until a number is found that is divisable
 by *"Number"* (default= 99) and by *"proof"* of the previous block.


``` {r, echo=TRUE}
 proof_of_work <- function(last_proof,Number=99)
   {
    proof <- last_proof + 1
  while (!(proof %% Number == 0 & proof %% last_proof == 0 ))
    {
    proof <- proof + 1
    }
 return(proof)
}
```

#### Function **"Nonce"**

In the case of BitCoin the PoW problem involves the problem of finding a number (called nonce)
that once added to the block is such that the corresponding block hash contains a certain amount
of leading zeros called difficulty (more specifically, Hashcash).
The average work that a miner needs to perform in order to find a valid nonce is exponential in the difficulty,
while one can verify the validity of the block by executing a single hash function.

``` {r, echo=TRUE}

Nonce <-  function(block,difficulty=1)
  {
  block$proof  <- 0
  hash <-   digest(block, "sha256")
  zero <-   paste(rep("0", difficulty), collapse="")
  while(substr(hash, 1, difficulty) != zero)
    {
    block$proof <- block$proof + 1
    hash <-  digest(block, "sha256")
    }
  block$proof
}
```

### Genesis Block

In order to create a new blockchain, first a so-called genesis block has to be built: this is provided by the function **block_gen**:

``` {r, echo= TRUE}
block_gen     <- function(data,prevHash="0")
{
 # data:     Data for the new chain
 # prevHash: Arbitrary string serving as previous hash

  block_genesis <-  list(index = 1,
                         timestamp = Sys.time(),
                         data = data,
                         previous_hash = prevHash ,
                         proof = 1)
  return(hash_block(block_genesis))
}
```
##### Example:
``` {r, echo= TRUE}
myData <- data.frame(x=rnorm(5),y=rpois(5,lambda=3))
GenBlock <- block_gen(data=list(myData))
print(GenBlock)
```



### Generating a new block

Generating a single new block as successor of a "previous block" is provided by the function **gen_new_block**:

```{r, echo=TRUE}
gen_new_block <- function(previous_block,data,PoW_func="proof_of_work",...)
{
# previous_block: Previous block.
# data:           Data for the new block.
# PoW_func:       String with the name Proof-of-Work function.
# ... :           Additional parameter(s) passed to the Proof-of-Work function.


  #Create new Block
  new_block <- list(index = previous_block$index + 1,
                    timestamp = Sys.time(),
                    data = data,
                    previous_hash = previous_block$new_hash)

  #Proof-of-Work
  if(PoW_func=="proof_of_work") 
      {new_proof <- proof_of_work(last_proof=previous_block$proof,...)}
  if(PoW_func=="Nonce")     new_proof <- Nonce(new_block,...)
  if(!(PoW_func %in% c("proof_of_work","Nonce"))) 
    {stop("This PoW-function is not implemented!")}
  new_block$proof  <- new_proof

  #Hash the new Block
  new_block_hashed <- hash_block(new_block)

  return(new_block_hashed)
}
```

##### Example:
``` {r, echo= TRUE}
NewData     <- list("Jordan",matrix(1:4,ncol=2),hist(rnorm(234),plot=FALSE),200)
prev_block  <- block_gen(data="Stephanie")
print(prev_block)
block2         <- gen_new_block(previous_block=prev_block,data=NewData[[1]]
                                ,PoW_func="proof_of_work",Number=33)
print(block2)
```



### Add new blocks to a chain
Adding blocks to an existing blockchain:

```{r, echo=TRUE}
add_blocks <- function(blockchain,n_add,new_data,PoW_func="proof_of_work",...)
{
#  blockchain: Blockchain to which new blocks are added.
#  n_add:      Number of blocks to be added.
#  new_data:   New data, list of length \code{n_add}.
#  PoW_func:   String with the name Proof-of-Work function.
#  ... :       Additional parameter(s) passed to the Proof-of-Work function.

  lengthChain    <- length(blockchain)
  if(!is.list(new_data)) stop("new_data must be in form of a list!")
  if(n_add > length(new_data)) stop("The number of new blocks must 
                                    not be greater then the number of new data.")
  previous_block <- blockchain[[lengthChain]]
  newChain       <- blockchain
  for (i in 1:n_add)
  {
    block_to_add <- gen_new_block(previous_block,data=new_data[[i]]
                                  ,PoW_func=PoW_func,...)
    newChain <- c(newChain,list(block_to_add))
    previous_block  <- block_to_add
  }
  newChain
}
```
##### Example:
```{r, echo=TRUE}
blockchain1  <- list(block_gen(data="Stephanie"))
print(blockchain1)

NewData      <- list("Jordan",matrix(1:4,ncol=2),hist(rnorm(234),plot=FALSE),200)
n            <- length(NewData)
blockchain2  <- add_blocks(blockchain=blockchain1, n_add=n,
                       new_data=NewData,PoW_func = "proof_of_work",Number=33)
```


### Creating a new chain

The function **new_chain** initializes a new blockchain based on a list containing data:

```{r, echo= TRUE}

new_chain <- function(data,PoW_func,...)
{
# data: List with the data for the new chain.
# PoW_func:   String with the name Proof-of-Work function.
# ... :       Additional parameter(s) passed to the Proof-of-Work function.

  if(!is.list(data)) stop("data must be in form of a list")
  n       <- length(data)
  if(n > 1)
  {
  X.0     <- list(block_gen(data[[1]],prevHash="0"))
   n.add   <- n-1
   Xdata   <- data[-1]
   X       <- add_blocks(X.0,n.add,Xdata)
  } else  X <- list(block_gen(data[[1]],prevHash="0"))
  X
  }
```
##### Example:
```{r, echo=TRUE}
MyData <- list(224,matrix(1:6,ncol=2),"Stephanie",c(220,230))
NewChain <- new_chain(data=MyData,PoW_func="Nonce",difficulty=2)
print(NewChain)
```

### Validating a chain

A simple validation of a block chain. The validation has two parts:

1. Checking that the encoding gives the correct hashes.
2. Checking that the parent field of a non-genesis block references to the previous block in the chain.

The function **validate_chain** takes as input a blockchain and  its output is a logical value: TRUE if the validation was successful. 

```{r, echo=TRUE}
validate_chain <- function(blockchain)
{
  for (i in 1:length(blockchain))
  {
    #print(paste("block:",i))
    block <- blockchain[[i]]
    hash <-  block$new_hash
    #print(hash)
    block$new_hash  <- NULL

    hash_expected <-  hash_block(block)$new_hash
    #print(hash_expected)
    if (hash != hash_expected) {
      return(FALSE)
    }
  }
  if (length(blockchain) >= 2)
  {
    for (i in 2:length(blockchain))
    {
      if (blockchain[[i]]$previous_hash != blockchain[[i-1]]$new_hash)
      {
        print(paste("Parent structure",(i-1)," ->",i,"is wrong."))
        return(FALSE)
      } else
        return(TRUE)
    }
  }
}
```
##### Example:
```{r, echo=TRUE}
MyData <- list(224,matrix(1:6,ncol=2),"Stephanie")
NewChain <- new_chain(MyData)
validate_chain(NewChain)
```

## Workflow

In order to create a new chain and to add new blocks, only the functions **new_chain** and **add_blocks** are needed; all the other functions are then internally called within these two functions.

### Example

First, we create a new chain whose data are give as:

```{r, echo=TRUE}
MyData <- list(224,matrix(1:6,ncol=2),"Stephanie",c(220,230))
newChain <- new_chain(data=MyData,PoW_func="Nonce",difficulty=2)
print(newChain)
```

In order to add new block(s) from new data, the function **add_blocks** is used:

```{r, echo=TRUE}
newData <- list(data.frame(x1=rnorm(5),Gender=c("f","f","m","f","m")),1:9)
updatedChain <- add_blocks(blockchain=newChain,n_add=2,new_data=newData,
                      PoW_func="Nonce",difficulty=2)
print(updatedChain)
```


